MCC Predictions

### 2018-04-24 stats/ex3_comparisons.Rmd
adding weak prior experiments

library(tidyverse)
library(knitr)
library(ggthemes)
```

@@

# check if someone did both experiments
x <- dat %>%
  filter(trial_type == "train") %>%
  group_by(id)%>%
  summarise(n = length(correct_inf)) %>%
  filter(n > 2)

# exclude those who dod both experiments
data <- dat %>%
  filter(id != x$id) %>%
  mutate(Change = ifelse(change =="same", "Same speaker", "Different speaker"),
         Alignment = ifelse(alignment == "congruent", "Congruent", "Incongruent"),
         Experiment = ifelse(experiment == "pref_inf", "Preference", "Novelty"))

@@

```{r comparisons to chance}
data %>%  
  filter(trial_type == "test") %>%
  group_by(change ,experiment,alignment, id) %>%
  summarise(correct_inf = mean(correct_inf)) %>%
  summarize(correct_inf = list(correct_inf)) %>%
  group_by(change ,experiment,alignment) %>%
  mutate(mean = mean(unlist(correct_inf)),
         stat = wilcox.exact(unlist(correct_inf), mu = 0.5)$statistic,
         p_value = wilcox.exact(unlist(correct_inf), mu = 0.5)$p.value) %>%
  select(change ,experiment,alignment,mean,stat,p_value)%>%
  knitr::kable(digits = 3)
```

@@

lm <- glmer(correct_inf ~ Experiment*Change*Alignment + (Change*Alignment|id) + (Change*Alignment|agent), 
              data = m.data, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

```{r plot models and data}
ggplot(pd, 
       aes(x = Change, y = mean, fill = Alignment)) +
  geom_bar(stat="identity", position = position_dodge(), color = 'black') + 
  geom_hline(yintercept = 0.5, lty=2)+
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper),position = position_dodge(width = 0.9))+
  labs(x=" ",y="Proportion Choosing More Informative")+
  facet_grid(Experiment ~ Model)+
  theme_few(base_size = 12) + 
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  ylim(0,1)
```
```{r plot comparing models}

@@

>>>>

library(tidyverse)
library(knitr)
library(ggthemes)
library(langcog)
```

@@

# check if someone did both experiments
# yes, two people
x <- dat %>%
  filter(trial_type == "train") %>%
  group_by(id)%>%
  summarise(n = length(correct_inf)) %>%
  filter(n > 2)

# exclude those who did both experiments
data <- dat %>%
  filter(!id %in% x$id) %>%
  filter(trial_type != "train") %>%
  mutate(Change = ifelse(change =="same", "Same speaker", "Different speaker"),
         Alignment = ifelse(alignment == "congruent", "Congruent", "Incongruent"),
         Experiment = ifelse(experiment == "pref_inf", "Preference", "Novelty"))

data %>%
  group_by(experiment,alignment,change) %>%
  summarise(n = length(id))
```

@@

```{r comparisons to chance}
library(exactRankTests)

data %>%  
  filter(trial_type == "test") %>%
  group_by(change ,experiment,alignment, id) %>%
  summarise(correct_inf = mean(correct_inf)) %>%
  summarize(correct_inf = list(correct_inf)) %>%
  group_by(change ,experiment,alignment) %>%
  mutate(mean = mean(unlist(correct_inf)),
         stat = wilcox.exact(unlist(correct_inf), mu = 0.5)$statistic,
         p_value = wilcox.exact(unlist(correct_inf), mu = 0.5)$p.value) %>%
  select(change ,experiment,alignment,mean,stat,p_value)%>%
  knitr::kable(digits = 3)
```

@@

lm <- glmer(correct_inf ~ Experiment*Change*Alignment + (Change*Alignment|id) + (Change+Alignment|agent), 
              data = m.data, family = binomial, 
              control=glmerControl(optimizer="bobyqa",optCtrl=list(maxfun=2e5)))

@@

```{r plot models and data}

ggplot(pd, 
       aes(x = Change, y = mean, fill = Alignment)) +
  geom_bar(stat="identity", position = position_dodge(), color = 'black') + 
  geom_hline(yintercept = 0.5, lty=2)+
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper),position = position_dodge(width = 0.9))+
  labs(x=" ",y="Proportion Choosing More Informative")+
  facet_grid(Experiment ~ Model)+
  theme_few(base_size = 12) + 
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1))+
  ylim(0,1)

```
```{r plot comparing models}

@@

```{r rsaUtils}
library(rwebppl)

rsaUtils <- '
var all_objects = [
{ shape: "triangle", id:1, location: 1},  
{ shape: "triangle", id:2, location: 2},
{ shape: "circle", id:1, location: 2}
]

var labels = ["dax","wug"]

var lexicon1 = function(utterance, obj){
utterance.label == "dax" ? obj.shape == "triangle" :
utterance.label == "wug" ? obj.shape == "circle" : 
true
}

var lexicon2 = function(utterance, obj){
utterance.label == "dax" ? obj.shape == "circle" :
utterance.label == "wug" ? obj.shape == "triangle" : 
true
}

var lexiconObjects = {
"dax = triangle": {
triangle: "dax", circle: "wug"
},
"dax = circle": {
triangle: "wug", circle: "dax"
},
}

var lexiconObject = {
"dax = triangle": lexicon1,
"dax = circle" : lexicon2
}

var point = function(utterance, obj){
return obj.location == utterance.point
}

var utterancePrior = function(obj, lexiconName){
var locationsWithShape = _.map(_.filter(all_objects, {shape: obj.shape}), "location")
var point = uniformDraw(locationsWithShape)
var label = lexiconObjects[lexiconName][obj.shape]
return {label: label, point: point}
}

var LexiconPrior = Categorical({vs: ["dax = triangle","dax = circle" ], ps: [1, 1]})
'
```

```{r rsaModel}
rsaModel <- 'var ObjectPrior = Categorical({vs: all_objects, ps: priorProbs })

var pragmaticListener = function(utterance){
Infer({method: "enumerate", model: function(){
var lexiconName = sample(LexiconPrior);
var obj = sample(ObjectPrior);
var S1 = speaker(obj, lexiconName);
observe(S1, utterance)
return {lexicon: lexiconName, obj: obj.shape}
}})
}

var speakerOptimality = 2.25;

var speaker = function(obj, lexiconName){
Infer({method: "enumerate", model: function(){
var utterance = utterancePrior(obj, lexiconName);
var L0 = literalListener(utterance);
 factor(speakerOptimality * L0.score(obj.shape))
return utterance
}})
}

var literalListener = function(utterance){
Infer({method: "enumerate", model: function(){
var lexiconName = sample(LexiconPrior); 
var lexicon = lexiconObject[lexiconName];
var obj = sample(ObjectPrior);
if ("label" in utterance) {
 var truthValue = lexicon(utterance, obj);
 condition(truthValue)
}
if (utterance.point) {
 var truthValuePoint = point(utterance, obj);
 condition(truthValuePoint)
}
return obj.shape 
}})
}

pragmaticListener({label: "dax", point: 2 })
'
```

```{r}
priorProbsFromR <- list(
  "pref_inf_same_congruent" =  c(0.0333333, 0.033333, 0.9666667),
  "pref_inf_same_incongruent" = c(0.9666667, 0.9666667, 0.0333333),
  "pref_inf_diff_congruent" =  c(0.3583333, 0.3583333, 0.6416667),
  "pref_inf_diff_incongruent" = c(0.6416667, 0.6416667, 0.3583333),
  "novel_inf_same_congruent" =  c(0.1666667, 0.1666667, 0.8333333),
  "novel_inf_same_incongruent" = c(0.8333333, 0.8333333, 0.1666667),
  "novel_inf_diff_congruent" =  c(0.4083333, 0.4083333, 0.5916667),
  "novel_inf_diff_incongruent" = c(0.5916667, 0.5916667, 0.4083333)
)

# get model predictions from webppl
wp.rs.all <- data.frame()
for (conditionName in names(priorProbsFromR)) {

  priorProbs <- priorProbsFromR[[conditionName]]
  rs.wp <- webppl(
    program_code = paste(rsaUtils, rsaModel, sep='\n'),
    data = priorProbs,
    data_var = "priorProbs"
  )
  wp.rs.all <- bind_rows(
    wp.rs.all, 
    rs.wp %>% 
      mutate(condition = conditionName,
             Model = "Full model")
  )
}

# common confition names to join by
con <- c("pref_inf_same_congruent",
  "pref_inf_same_incongruent",
  "pref_inf_diff_congruent",
  "pref_inf_diff_incongruent",
  "novel_inf_same_congruent",
  "novel_inf_same_incongruent",
  "novel_inf_diff_congruent",
  "novel_inf_diff_incongruent")

#Model predictions from prior only model
prior_only <- predictions %>%
  filter(Model == "Prior only")%>%
  mutate(condition = con)

#Model predictions from no prior model
no_prior <- predictions %>%
  filter(Model == "No prior")%>%
  mutate(condition = con)

# summarizing the data
data.summary <- data %>%
  mutate(condition = paste(experiment,change, alignment, sep = "_")) %>%
  group_by(condition) %>%
  summarize(n= n(),
            k = sum(correct_inf))

# joining model predictions and data
model_comparison <- rbind(
  left_join(
  wp.rs.all %>% filter(obj == "circle")%>%mutate(exp = substr(condition,1,4)) %>%select(condition,exp,Model,prob),
  data.summary),
  left_join(
  no_prior%>%mutate(prob = corr, exp = substr(condition,1,4))%>%select(condition,exp,Model,prob),
  data.summary),
  left_join(
  prior_only%>%mutate(prob = corr, exp = substr(condition,1,4))%>%select(condition,exp,Model,prob),
  data.summary)
  )

# calculate how well model predicts data
comp_model<- model_comparison%>%
  mutate(p_data_given_model = dbinom(size = n, p = prob, x = k, log = T)) %>%
  group_by(Model) %>%
  summarize(log_p_data_given_model = sum(p_data_given_model))

comp_model

## log p (data | model)
## bayes factor_1,2 = p (data | model_1) / p (data | model_2)
## log bf = log p (data | model_1) - log p (data | model_2)

# Bayes factors full vs other models
bf_full_no_prior <- exp((comp_model[1,2] - comp_model[2,2]))
bf_full_no_prior

bf_full_prior_only <- exp((comp_model[1,2] - comp_model[3,2]))
bf_full_prior_only

# model comparison by experiment
comp_model_by_exp<- model_comparison%>%
  mutate(p_data_given_model = dbinom(size = n, p = prob, x = k, log = T)) %>%
  group_by(Model,exp) %>%
  summarize(log_p_data_given_model = sum(p_data_given_model))

# Bayes factors full - prior only
pref_bf_full_prior_only <- exp(comp_model_by_exp[2,3] - comp_model_by_exp[6,3])
pref_bf_full_prior_only

novel_bf_full_prior_only <- exp(comp_model_by_exp[1,3] - comp_model_by_exp[5,3])
novel_bf_full_prior_only
